---
title: "R: Basic Operations"
output: html_notebook
---

# R: Basic Operations

## Comments

Comments in R are denoted by `#`

```{r}
# This is a comment
###Nazrin

#There is no built-in way to do multi-line comments.

## Mathematical Operations

#Mathematical operations in R are handled using the standard 4 symbols `+, -, /, and *` . Operations are calculated using PEMDAS.

### Addition
5+1
# Addition

```

### Subtraction

6-2
### Multiplication

3*5
6*2

### Division
25/5
24/6

## Variable Assignment

Variables in R are traditionally assigned using an arrow sign `<-` , made using a less than carat and a dash. It is valid to use `=` , but not recommended.

x<-5
y<-4
z<-x+y

Variable names can not include and spaces or symbols other than the underscore `_` . They cannot begin with numbers, but can include numbers after the first character.
mynumber_2<-7
variable_8<-3

## Data Types

R has the following data types: numeric, integer, character, logical, and complex.

### Numeric

Decimal values in R are numerics. Numerics are the default data type for numbers.

a<-6.75
b<-34.76

Note: Numerics are not integers. You can confirm this with both a and b above using the `is.integer()` function.

is.integer(a)
is.integer(b)
### Integer

To create an integer in R we can use the `as.integer()` function.
c<-as.integer(3.2)
is.integer(c)

For shorthand, we can also define an integer by writing the desired integer value, followed by a capital L.

d<-9L
is.integer(d)

Integers must be whole numbers, if passed a floating value, the number will be rounded to an integer.
 
 b<-as.integer(5.6)
 
This also works with a string.

t<-as.integer("8")

We would use an integer data-type when we want to represent a whole integer, and don't want to accept any other variations of numbers. The numeric data-type is less strict, it allows us to switch between different number types like integers and floats dynamically.

### Character

A **character** type is used to represent strings in R. Characters can be denoted by writing within quotes.

```{r}
# Defining a character object

my_string <- "Hello, Nazrin!"

#### Concatenating character objects

Character objects can be concatenated into a single variable using the `paste` f
greeting <- paste("Hello", "Nazrin")


#### Converting into character objects

We can convert other objects into character objects using the `as.character()` f

num <- 123
char_num <- as.character(num)

# Note that class returns the data type of our object

class(char_num)  # "character"


#### Printing character types


You can print out hardcoded character objects with the `print` function

print("This is a string in R")

You can print out formatted character objects with the `sprintf` function. You can use the following symbols to represent formatted parts of the code, each should be preceded by a `%` symbol:

-   d and i - Integer values

-   s - Character string

-   o - Octal

-   x and X - Hexadecimal

-   f - Double precision value (For fixed point representation, for example: 21.234)

-   e and E - Double precision value (For exponential representation, for example: 2.1234e+01)

name <- "Nazrin"
sprintf("My name is %s.", name)

#### Return a substring

You can get a substring from a character object using the `substr` function

substr("OpenAI", 1, 4)  # "Open"

#### Replacing characters in a string

You can replace characters in a string using the `sub` function

sub("Nazrin", "World", "Hello Nazrin")  # "Hello World"

### Logical


A logical value is commonly referred to as a boolean value. It is the result of a comparison between two variables. It is represented as `TRUE` or `FALSE.`

h <- TRUE
i <- FALSE

#### Logical Operators

The standard logical operations are:

-   `&` - And

-   `|` - Or

-   `!` - Not

j <- h & i  # FALSE
k <- h | i  # TRUE
l <- !h     # FALSE

### Complex

A complex object in R is created using the symbol `i` . The use of the i symbol is in reference to imaginary numbers.

m <- 2 + 3i


Re(m)  # 2
Im(m)  # 3

## Comparison Operators

You can use the following operators to compare two variables and return a logical object:

-   `<` Less than

-   `>` Greater than

-   `<=` Less than or equal to

-   `>=` Greater than or equal to

-   `==` Is equal to

-   `!=` Is NOT equal to

n <- 5
o <- 10
n < o   # TRUE
n > o   # FALSE
n <= o  # TRUE
n >= o  # FALSE
n == o  # FALSE
n != o  # TRUE
 

## Homework

1.  Create two strings with your first and last name, respectively. Then concatenate them, and print the new value as "My name is \<your newly created concatenated variable will go here\>"

first_name <- "Nazrin"
last_name <- "Lastname"
full_name <- paste(first_name, last_name)
print(paste("My name is", full_name))

2.  What values can a BOOLEAN contain?

  A BOOLEAN in R can contain TRUE or FALSE . 
  
  